---
title: "Assignment 4  - MIS64060"
author: "Tim Pagliari tpagliar@kent.edu"
date: "10/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
dir.create("images")
options(scipen=999)
library(dplyr)
library(tidyverse)
library(fastDummies)
library(factoextra) # clustering algorithms & visualization
library(ISLR)
library(caret)
```

## Assignment details:
The dataset on American College and University Rankings contains information on 1302 American colleges and
universities offering an undergraduate program. For each university, there are 17 measurements, including
continuous measurements (such as tuition and graduation rate) and categorical measurements (such as location by
state and whether it is a private or public school). Note that many records are missing some measurements.
<ul>
<li> Remove all records with missing measurements from the dataset.</li>
<li> For all the continuous measurements, run K-Means clustering. Make sure to normalize the
measurements. How many clusters seem reasonable for describing these data? What was
your optimal K?</li>
<li> Compare the summary statistics for each cluster and describe each cluster in this context
(e.g., “Universities with high tuition, low acceptance rate...”).</li>
<li> Use the categorical measurements that were not used in the analysis (State and
Private/Public) to characterize the different clusters. Is there any relationship between the
clusters and the categorical information?</li>
<li> What other external information can explain the contents of some or all of these clusters?</li>
<li> Consider Tufts University, which is missing some information. Compute the Euclidean
distance of this record from each of the clusters that you found above (using only the
measurements that you have). Which cluster is it closest to? Impute the missing values for
Tufts by taking the average of the cluster on those measurements.</li>
</ul>

## Import and tidy the data

```{r}
UniversitiesFull <- read.csv("/Users/tpagliar/GIT/MIS64060-tpagliar/Universities.csv", stringsAsFactors = TRUE)
UniversitiesFull <- dummy_columns(UniversitiesFull, select_columns = 'Public..1...Private..2.')
colnames(UniversitiesFull) <- c("Name","State","PublicPrivate","Applied","Accepted","Enrolled","Top10","Top25","FullTimeUG","PartTimeUG","InStateTuition","OutStateTuition","Room","Board","Fees","EstBookCost","EstPersCost","FacWPHD","StudFacRatio","GradRate","Public","Private")
Universities <- UniversitiesFull
summary(Universities)
Universities$PublicPrivate <- as.factor(Universities$PublicPrivate)
Universities$AcceptRate<-((Universities$Accepted/Universities$Applied)*100) #I'll use these to simplify the contribution to the selectivity and yield of each school.
Universities$EnrolledRate<-((Universities$Enrolled/Universities$Accepted)*100)#I'll use these to simplify the contribution to the selectivity and yield of each school. 
Universities$AcceptRate<-((Universities$Accepted/Universities$Applied)*100)
Universities$EnrolledRate<-((Universities$Enrolled/Universities$Accepted)*100)
colMeans(is.na(Universities)*100) #we see there are many incomplete records, we'll remove them from our dataset
Universities <- Universities[complete.cases(Universities),] #Include only complete data
summary(Universities) #we see that some schools have data that is a heavy outlier, ex: a 118% graduation rate or 103% faculty with PhDs. We'll filter these 'erroneous' rows below:
Universities <- filter(Universities, (GradRate <= 100) & (FacWPHD <= 100))#we see that some schools have data that is a heavy outlier, ex: a 118% graduation rate or 103% faculty with PhDs. 
summary(Universities)

```

## Normalizing Data
```{r}
#columnsare <- c(3,)
ScaledValues <- scale(Universities[,4:20], center=TRUE,scale=TRUE)
UniversitiesScaled <- Universities
UniversitiesScaled[,4:20]<- ScaledValues
summary(UniversitiesScaled)
```

## Clustering
We'll start with 5 clusters as a base, then will do some analysis to determine the optimal number of centers.

```{r}
k5 <- kmeans(ScaledValues, centers = 5, nstart = 25) # k = 5, number of restarts = 25
k5$centers # output the centers
k5$size # Number of schools in each cluster
k5$cluster[120] # Identify the cluster of the 120th observation as an example
fviz_cluster(k5, data = ScaledValues) # Visualize the output
```

## Optimal Clustering
```{r}
set.seed(321)
fviz_nbclust(ScaledValues, kmeans, method = "wss")
#It looks like 3 or 4 is the optimal n-cluster here....
fviz_nbclust(ScaledValues, kmeans, method = "silhouette")
#We'll run kmeans with 3 centers as recommended by the silhouette
k3 <- kmeans(ScaledValues, centers = 3, nstart = 25) # k = 4, number of restarts = 25
k3$centers # output the centers
k3$size # Number of schools in each cluster
k3$cluster[120] # Identify the cluster of the 120th observation as an example
kmeansplot = fviz_cluster(k3, data = ScaledValues) # Visualize the output
kmeansplot
```

<ul>
<li>Cluster 1 has low enrollment and lower student profile, medium student:faculty ratio and few faculty with PhDs </li>
<li>Cluster 2 has high performing students, high selectivity and a relatively high cost. It also has a low student:faculty ratio and many PhD faculty. Most importantly, it has a very high Graduation Rate.</li>
<li>Cluster 3 has high enrollment, low selectivity, low tuition but high costs, large class sizes and medium performance by its students. </li>
</ul>

```{r}
#now lets check the categorical variables
Universities$ClusterNum <- as.factor(k3$cluster)
with(Universities, table(ClusterNum,PublicPrivate))
with(Universities, table(ClusterNum,State))
```
We can see that group 2 - the high performers - are almost entirely private (98%); and that group 3 - the large schools - are 89% public. A correlation by state is not obvious.

## Tufts University
```{r}
UniversitiesTufts<-Universities
UniversitiesTufts[nrow(UniversitiesTufts)+1,]<-UniversitiesFull[which(UniversitiesFull$Name == "Tufts University"),]
#we'll normalize it as before

ScaledValuesT <- scale(UniversitiesTufts[,4:20], center=TRUE,scale=TRUE)
UniversitiesScaledT <- UniversitiesTufts
UniversitiesScaledT[,4:20]<- ScaledValuesT
ScaledValuesT[which(UniversitiesTufts$Name == "Tufts University"),]
clustercenters<-k3$centers
distance1 <- rbind(ScaledValuesT[which(UniversitiesTufts$Name == "Tufts University"),],clustercenters[1,])
distance2 <- rbind(ScaledValuesT[which(UniversitiesTufts$Name == "Tufts University"),],k3$centers[2,])
distance3 <- rbind(ScaledValuesT[which(UniversitiesTufts$Name == "Tufts University"),],k3$centers[3,])
dist(distance1,method="euclidean") #distance from Tufts to Cluster 1 center
dist(distance2,method="euclidean") #distance from Tufts to Cluster 2 center
dist(distance3,method="euclidean") #distance from Tufts to Cluster 3 center
```
Tufts is closest (at at 2.73 units) to <b>Cluster 2 , the high performer cluster</b>. Tufts is missing their Part-Time UG, so we can impute this from the cluster 2 average (-0.3144979).
