---
title: "tpagliar_1.rmd"
author: "Tim Pagliari tpagliar@kent.edu"
date: "9/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE)
```

#### Below is Tim Pagliari (tpagliar@kent.edu)'s submission for Assignment 1 in MIS-64060-002-202080.  
Data used is from the CScoreCard dataset, which is the Most Recent Institution Level Data from the Department of Education's College Scorecard. Retrieved from: https://collegescorecard.ed.gov/data/; I've removed some of the later columns as the original file is greater than 100MB and I didn't try to push it until i had completed the assignment.

`r CScoreCard <- read.csv("~/GIT/MIS64060-tpagliar/CScoreCard.csv")`
The MaxDegree variable is categorical and references the highest degree level awarded by the institution. 
```{r}
count_Rows <- nrow(CScoreCard) #First, I'll count the number of rows (or number of institutions present)
var_MaxDegree <- CScoreCard$HIGHDEG
count_BSplus <- sum(var_MaxDegree>=3) #Find those that award Baccalaureate Degrees or Higher
count_MSplus <- sum(var_MaxDegree>=4) #Find those that award Master's Degrees or Higher
```
There are **`r count_Rows`** schools in the dataset, using the MaxDegree information we know that **`r count_BSplus`** schools award Bachelor's degrees or higher, and **`r count_MSplus`** schools award Master's degrees or higher.

The **SAT_AVG** variable is qualitative, and demonstrates the average SAT score of students enrolled at the institution.
```{r}
sat<-as.numeric(CScoreCard$SAT_AVG, na.rm=TRUE)
var_maxsat <- max(sat,na.rm=TRUE) #is the maximum SAT average among schools who reported.
SchoolWMAXSAT <- CScoreCard[which.max(sat),4] #This score belongs to:
sqrt_sat <- sqrt(sat) #An example transformation of the SAT scores into the square root of the SAT scores
act_matmid<-as.numeric(CScoreCard$ACTMTMID,na.rm=TRUE)
```
Using this variable, we know that **`r var_maxsat`** is the maximum SAT average among schools who reported this data, and that score belongs to **`r SchoolWMAXSAT`**. If we wanted to do a transformation on that score, the **sqrt_sat** is the square root of the SAT scores.

#### Plots of the SAT vs ACT midpoints and a histogram of the admit rate frequencies are below.
```{r}
plot(act_matmid,sat)#Here is a plot of the SAT scores (y axis) vs the midpoint ACT Math score (x axis)
hist(as.numeric(CScoreCard$ADM_RATE))#and here is a histogram of the admit rate frequencies. There seem to be fewer selective institutions in this list.
sqrt_sat <- sqrt(sat) #An example transformation of the SAT scores into the square root of the SAT scores
```


